On exponential sums arising from the classical doubling method

### 2025-04-10 main.tex
Explaining the cocharacter notation

\newcommand{\GaussSumScalar}[2]{\mathrm{G}\left(#1, #2\right)}
\newcommand{\dblJacobiSumScalar}[2]{\mathrm{J}_{\pm}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\posDblJacobiSumScalar}[2]{\mathrm{J}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\negDblJacobiSumScalar}[2]{\mathrm{J}_{-}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblVirtualJacobiSumScalar}[2]{\mathrm{j}_{\pm}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblPosVirtualJacobiSumScalar}[2]{\mathrm{j}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblNegVirtualJacobiSumScalar}[2]{\mathrm{j}_{-}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\posDblVirtualJacobiSumScalar}[2]{\mathrm{j}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblGammaFactor}[3]{\Gamma^{\mathrm{dbl}}\left(#1 \times #2, #3\right)}
\newcommand{\dblGammaFactorSpace}[4]{\Gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\dblLanglandsGammaFactorSpace}[4]{\gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\GaussSumCharacter}[3]{\tau\left(#1 \times #2, #3\right)}
\newcommand{\GaussSumTorusCharacter}[4]{\tau_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\ladicnumbers}{\algebraicClosure{\mathbb{Q}_{\ell}}}
\newcommand{\IsometryGroup}{\mathrm{Isom}}
\newcommand{\lieAlgebra}{\mathfrak{g}}
\newcommand{\DeligneLusztigInduction}[2]{\mathrm{R}_{#1}^{#2}}
\newcommand{\algebraicGroup}[1]{\boldsymbol{\mathrm{#1}}}
\newcommand{\LusztigSeries}[2]{\mathcal{E}\left(#1, (#2)\right)}
\newcommand{\characteristicPolynomial}{\operatorname{CharPoly}}
\newcommand{\DualFrobeniusFixedPoints}[2][\Frobenius^{\ast}]{\algebraicGroup{#2}^{\ast #1}}
\newcommand{\FrobeniusFixedPoints}[2][\Frobenius]{\algebraicGroup{#2}^{#1}}
\newcommand{\CharacterLattice}[1]{X^{\ast}\left(#1\right)}
\newcommand{\CocharacterLattice}[1]{X_{\ast}\left(#1\right)}
\newcommand{\SymmetricGroup}{\mathfrak{S}}
\newcommand{\Kernel}{\operatorname{Ker}}
\newcommand{\RTTheta}[2]{R_{#1, #2}}
\newcommand{\RTGTheta}[3]{R_{#1}^{#2}(#3)}
\newcommand{\RTThetaVirtual}[3]{\mathcal{R}_{#1}^{#2}(#3)}

@@

    \exp(\langle x_{T,\theta},y\rangle) = \theta((N_{T(\finiteField_m):T(\finiteField)}(y(\zeta_m))).

@@

@@

    \item ($n$-cycle) $\algebraicGroup{T}(\finiteField) \cong \finiteField_n^\times$ with $$\FieldNorm{\algebraicGroup{T}(\finiteField_{mn})}{\algebraicGroup{T}(\finiteField)}(y_1,y_2 \hdots y_n)(\zeta_{mn}) = (\zeta_n)^{y_1+qy_2+q^2y_3 \hdots +q^{n-1}y_n}$$
    \item ($n$-cycle with one negative) $\algebraicGroup{T}(\finiteField) \cong \{x \in \finiteField_{2n}^\times: \FieldNorm{2n}{n}(x) = 1\}$ with $$\FieldNorm{\algebraicGroup{T}(\finiteField_{2mn})}{\algebraicGroup{T}(\finiteField)}(y_1,y_2 \hdots y_n)(\zeta_{2mn}) = (\beta_n)^{y_1+qy_2+q^2y_3 \hdots +q^{n-1}y_n}$$

@@

    \item ($(-1) \cdot n$-cycle, $n$ even) $\algebraicGroup{T}(\finiteField) \cong \finiteField_n^\times$ with $$\FieldNorm{\algebraicGroup{T}(\finiteField_{mn})}{\algebraicGroup{T}(\finiteField)}(y_1,y_2 \hdots y_n)(\zeta_{mn}) = (\zeta_n)^{y_1-qy_2+q^2y_3-q^3y_4 \hdots -q^{n-1}y_n}$$
    \item ($(-1) \cdot n$-cycle, $n$ odd) $\algebraicGroup{T}(\finiteField) \cong \{x \in \finiteField_{2n}^\times: \FieldNorm{2n}{n}(x) = 1\}$ with $$\FieldNorm{\algebraicGroup{T}(\finiteField_{2mn})}{\algebraicGroup{T}(\finiteField)}(y_1,y_2 \hdots y_n)(\zeta_{2mn}) = (\beta_n)^{y_1-qy_2+q^2y_3-q^3y_4 \hdots +q^{n-1}y_n}$$

@@

Now the results follow from the equalities

>>>>

\newcommand{\GaussSumScalar}[2]{\mathrm{G}\left(#1, #2\right)}
\newcommand{\dblJacobiSumScalar}[2]{\mathrm{J}_{\pm}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\posDblJacobiSumScalar}[2]{\mathrm{J}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\negDblJacobiSumScalar}[2]{\mathrm{J}_{-}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblVirtualJacobiSumScalar}[2]{\mathrm{j}_{\pm}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblPosVirtualJacobiSumScalar}[2]{\mathrm{j}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblNegVirtualJacobiSumScalar}[2]{\mathrm{j}_{-}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\posDblVirtualJacobiSumScalar}[2]{\mathrm{j}_{+}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblGammaFactor}[3]{\Gamma^{\mathrm{dbl}}\left(#1 \times #2, #3\right)}
\newcommand{\dblGammaFactorSpace}[4]{\Gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\dblLanglandsGammaFactorSpace}[4]{\gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\GaussSumCharacter}[3]{\tau\left(#1 \times #2, #3\right)}
\newcommand{\GaussSumTorusCharacter}[4]{\tau_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\ladicnumbers}{\algebraicClosure{\mathbb{Q}_{\ell}}}
\newcommand{\IsometryGroup}{\mathrm{Isom}}
\newcommand{\lieAlgebra}{\mathfrak{g}}
\newcommand{\DeligneLusztigInduction}[2]{\mathrm{R}_{#1}^{#2}}
\newcommand{\algebraicGroup}[1]{\boldsymbol{\mathrm{#1}}}
\newcommand{\LusztigSeries}[2]{\mathcal{E}\left(#1, (#2)\right)}
\newcommand{\characteristicPolynomial}{\operatorname{CharPoly}}
\newcommand{\DualFrobeniusFixedPoints}[2][\Frobenius^{\ast}]{\algebraicGroup{#2}^{\ast #1}}
\newcommand{\FrobeniusFixedPoints}[2][\Frobenius]{\algebraicGroup{#2}^{#1}}
\newcommand{\CharacterLattice}[1]{X^{\ast}\left(#1\right)}
\newcommand{\CocharacterLattice}[1]{X_{\ast}\left(#1\right)}
\newcommand{\SymmetricGroup}{\mathfrak{S}}
\newcommand{\Kernel}{\operatorname{Ker}}
\newcommand{\RTTheta}[2]{R_{#1, #2}}
\newcommand{\RTGTheta}[3]{R_{#1}^{#2}(#3)}
\newcommand{\RTThetaVirtual}[3]{\mathcal{R}_{#1}^{#2}(#3)}
\newcommand{\Cocharacter}{\mathrm{cochar}}

@@

    \exp\left(\innerproduct{x_{T,\theta}}{y}\right) = \theta\left(\aFieldNorm_{\algebraicGroup{T}(\finiteField_m):\algebraicGroup{T}(\finiteField)}\left(y(\zeta_m\right)\right).

@@

In the following computations, we will choose $m$ to be sufficiently large so such that the torus $\algebraicGroup{T}\left(\finiteField_m\right)$ is isomorphic to $\left(\multiplicativegroup{\finiteFieldExtension{m}}\right)^N$. We will then identify the cocharacter lattice $\CocharacterLattice{\algebraicGroup{T}}$ with $\zIntegers^N$ by the map $\Cocharacter \colon \zIntegers^N \to \CocharacterLattice{\algebraicGroup{T}}$ given by $\left(y_1,\dots,y_N\right) \mapsto \left(\Cocharacter_{(y_1,\dots,y_N)}\left(t\right) = \left(t^{y_1},\dots,t^{y_N}\right)\right)$.

@@

    \item (cycle of length $n$) $\algebraicGroup{T}(\finiteField) \cong \finiteField_n^\times$ and $\algebraicGroup{T}\left(\finiteFieldExtension{nm}\right) \cong \left(\multiplicativegroup{\finiteFieldExtension{m}}\right)^n$ with $$\FieldNorm{\algebraicGroup{T}(\finiteField_{mn})}{\algebraicGroup{T}(\finiteField)}\left(\Cocharacter_{(y_1,\dots,y_n)}(\zeta_{mn})\right) = \prod_{i=0}^{n-1} \left(\zeta_n\right)^{q^{i} y_{i+1}}.$$
    \item (cycle of length $n$ with one negative sign) $\algebraicGroup{T}(\finiteField) \cong \NormOneGroup{2n}$ and $\algebraicGroup{T}\left(\finiteFieldExtension{2mn}\right) \cong \left(\multiplicativegroup{\finiteFieldExtension{2mn}}\right)^{n}$ with $$\FieldNorm{\algebraicGroup{T}(\finiteField_{2mn})}{\algebraicGroup{T}(\finiteField)}\left(\Cocharacter_{(y_1,\dots, y_n)}\left(\zeta_{2mn}\right)\right) = \prod_{i=0}^{n-1}(\beta_n)^{q^{i} y_{i+1}},$$

@@

    \item (cycle of length $n$ with $n$ negative signs, $n$ even) $\algebraicGroup{T}(\finiteField) \cong \finiteField_n^\times$ and $\algebraicGroup{T}\left(\finiteFieldExtension{mn}\right) \cong \left(\multiplicativegroup{\finiteFieldExtension{nm}}\right)^n$ with $$\FieldNorm{\algebraicGroup{T}(\finiteField_{mn})}{\algebraicGroup{T}(\finiteField)}\left(\Cocharacter_{(y_1,\dots,y_n)}(\zeta_{mn})\right) = \prod_{i=0}^{n-1} (\zeta_n)^{\left(-1\right)^{i} q^{i} y_{i+1}}.$$
    \item (cycle of length $n$ with $n$ negative signs, $n$ odd) $\algebraicGroup{T}(\finiteField) \cong \NormOneGroup{2n}$ and $\algebraicGroup{T}\left(\finiteFieldExtension{2mn}\right) \cong \left(\multiplicativegroup{\finiteFieldExtension{2mn}}\right)^{n}$ with $$\FieldNorm{\algebraicGroup{T}(\finiteField_{2mn})}{\algebraicGroup{T}(\finiteField)}\left(\Cocharacter_{(y_1,\dots, y_n)}\left(\zeta_{2mn}\right)\right) = \prod_{i=0}^{n-1} (\beta_n)^{\left(-1\right)^{i}q^{i}y_{i+1}},$$

@@

The results now follow from the equalities
